Motional Ground-State Cooling Outside the Lamb-Dicke Regime

### 2017-07-10 paper.tex
Axis explained

\begin{figure*}
  \includegraphics[height=5.2cm]{imgs/Na_RSC_schematic.pdf}
  \includegraphics[height=4.5cm]{sequence.pdf}
  \caption{(A) Energy levels and schematic of Raman sideband cooling.
    The Raman transitions have a one photon detuning $\Delta=25$ GHz from the D2 line.
    We use D1 light with $\sigma^-$ polarization to repump atoms out of $|F=2,m_F=-1\rangle$
    state to minimize heating on the atom in $|F=2,m_F=-2\rangle$ state
    (B) Geometry and polarizations of the Raman and optical pumping beams relative to the
    optical tweezer and bias magnetic field.
    (C) Schematic of the cooling sequence. The tweezer switches at 3 MHz to
    reduce light shifts during optical pumping. Each cooling cycle consists of $8$ pulses.
    The four axial pulses are addressing two neighboring cooling orders.
    The two pulses in each radial directions are either addressing two neighboring cooling orders
    or having different length on the first order when most of the population are below $n=3$
    towards then end of the cooling sequence.
    \label{f-setup}}
\end{figure*}
\begin{figure}[b]
  \includegraphics[width=8.5cm]{imgs/fig2_raman_op.pdf}
  \caption{Matrix elements and heating as a function of motional state.
    The range plotted covers $95\%$ of the initial thermal distribution.
    (A) Matrix elements for Raman transition in the axial direction showing deviation from
    $\sqrt{n}$ scaling and multiple minimums for different sideband orders.
    (B) Average change in motional state after the optical pumping step for all three axis.
    Due to the large Lamb-Dicke parameter,
    there is a high probability of $n$ changing especially in the axial direction.
    \label{f-ld}}
\end{figure}
\begin{figure*}
  \includegraphics[height=4.2cm]{imgs/spectrum_r.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_r3_0.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_r3_p1.pdf}
  \caption{(A) Radial Raman sideband spectrum of first order heating, first order cooling and
    second order cooling before and after Raman sideband cooling.
    (B,C) Rabi flopping on axis 3 (B) carrier and (C) first order heating sideband
    before (red) and after (blue) cooling.
    Solid lines in (B) and (C) are theoretical calculation of the Rabi flopping.
    The blue lines corresponds to a ground state probability of $93\%$ after cooling and
    the red lines corresponds to a thermal distribution of $70$ $\mu$K before cooling.
    \label{f-radial}}
\end{figure*}
\begin{figure*}
  \includegraphics[height=4.2cm]{imgs/spectrum_a1.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_a1_0.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_a1_p1.pdf}
  \caption{(A) Axial Raman sideband spectrum from first order heating to eighth order cooling
    before and after Raman sideband cooling.
    The data for the second and higher orders of cooling sidebands are taken with 150 $\mu$s
    pulse time and the rest are taken with 125 $\mu$s pulse time.
    (B,C) Rabi flopping on axial (B) carrier and (C) first order heating sideband
    before (red) and after (blue) cooling.
    Solid lines in (B) and (C) are theoretical calculation of the Rabi flopping.
    The blue and green lines corresponds to a ground state probability of $92\%$ after cooling and
    the red lines corresponds to a thermal distribution of $70$ $\mu$K before cooling.
    The blue lines does not take into account the effect of decoherence due to resonance
    fluctuation. By comparing it to the green line in (C), which includes a $3$ kHz fluctuation,
    we can clearly see that this effect is the strongest for the after cooling data on
    the axial heating sideband where the Rabi frequency is the lowest.
    \label{f-axial}}
\end{figure*}

@@

Our final cooling results are shown in Fig.~\ref{f-radial} and \ref{f-axial}.
In total, $1000$ cooling pulses (for $100$ ms) are applied along three axes with
cooling begins on the radial second order to remove most energy initially.
To characterize the single atom thermal occupation before and after cooling,
we perform Raman sideband thermometry before and after cooling\fxnote{add citation?}.
\fxnote{How is the thermometry done? pulse duration etc.}
For the more tightly confined radial directions (non-degenerate),
we observe clear first-order heating, first-order cooling,
and second-order cooling sidebands before RSC as is shown in red in Figure \ref{f-radial}A.
After cooling, the first and second order cooling sidebands on both radial axes are suppressed.
Given the absence of the second order cooling sideband, we perform sideband thermometry
by taking the ratio of the peaks of first-order cooling and heating sidebands to measure $\bar n / (\bar n + 1)$ \fxnote{cite}.
We found $\{\bar n =,\bar n =\}$, which corresponds to a ground state population of $90(2)\%$
and $94(3)\%$ along the two radial directions.
It is important to note that the simple sideband thermometry formula assumes the coupling strength
to be proportional to $\sqrt{n}$. Outside of the Lamb-Dickes regime,
the coupling strength deviates from this simple scaling rule rapidly
as shown in figure \ref{f-ld}A. Since the Raman spectroscopy is performed
with a relatively large Raman Lamb-Dickes parameters $\eta^R$,
we therefore use an independent measurement of Rabi flopping on the carrier (Fig. \ref{f-radial}B)
and first-order heating(Fig. \ref{f-radial}C) sidebands
to verify the ground-state population~\cite{Meekhof1996}.
The data is fitted using the same Monte-Carlo simulation we used to simulate the cooling
which yields $89(2)\%$ and $92(2)\%$ and show good agreement with
the population extracted from the sideband thermometry.
We also extracted the initial temperature before cooling
from the motional-state decohered Rabi flopping and obtained \fxnote{70(?)} $\mu$K.

>>>>

\begin{figure*}
  \includegraphics[height=5.2cm]{imgs/Na_RSC_schematic.pdf}
  \includegraphics[height=4.5cm]{sequence.pdf}
  \caption{(A) Energy levels and schematic of Raman sideband cooling.
    The Raman transitions have a one photon detuning $\Delta=25$ GHz from the D2 line.
    We use D1 light with $\sigma^-$ polarization to repump atoms out of $|F=2,m_F=-1\rangle$
    state to minimize heating on the atom in $|F=2,m_F=-2\rangle$ state
    (B) Geometry and polarizations of the Raman and optical pumping beams relative to the
    optical tweezer and bias magnetic field.
    (C) Schematic of the cooling sequence. The tweezer switches at 3 MHz to
    reduce light shifts during optical pumping. Each cooling cycle consists of $8$ pulses.
    The four axial pulses are addressing two neighboring cooling orders.
    The two pulses in each radial directions are either addressing two neighboring cooling orders
    or having different length on the first order when most of the population are below $n=3$
    towards then end of the cooling sequence.
    \label{f-setup}}
\end{figure*}
\begin{figure}[b]
  \includegraphics[width=8.5cm]{imgs/fig2_raman_op.pdf}
  \caption{Matrix elements and heating as a function of motional state.
    The range plotted covers $95\%$ of the initial thermal distribution.
    (A) Matrix elements for Raman transition in the axial direction showing deviation from
    $\sqrt{n}$ scaling and multiple minimums for different sideband orders.
    ($k$ is the difference in wave vector of the Raman beams,
    $r$ is the coordinate in axial direction).
    (B) Average change in motional state after the optical pumping step for all three axis.
    Due to the large Lamb-Dicke parameter,
    there is a high probability of $n$ changing especially in the axial direction.
    \label{f-ld}}
\end{figure}
\begin{figure*}
  \includegraphics[height=4.2cm]{imgs/spectrum_r.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_r3_0.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_r3_p1.pdf}
  \caption{(A) Radial Raman sideband spectrum of first order heating, first order cooling and
    second order cooling before and after Raman sideband cooling.
    (B,C) Rabi flopping on axis 3 (B) carrier and (C) first order heating sideband
    before (red) and after (blue) cooling.
    Solid lines in (B) and (C) are theoretical calculation of the Rabi flopping.
    The blue lines corresponds to a ground state probability of $93\%$ after cooling and
    the red lines corresponds to a thermal distribution of $70$ $\mu$K before cooling.
    \label{f-radial}}
\end{figure*}
\begin{figure*}
  \includegraphics[height=4.2cm]{imgs/spectrum_a1.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_a1_0.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_a1_p1.pdf}
  \caption{(A) Axial Raman sideband spectrum from first order heating to eighth order cooling
    before and after Raman sideband cooling.
    The data for the second and higher orders of cooling sidebands are taken with 150 $\mu$s
    pulse time and the rest are taken with 125 $\mu$s pulse time.
    (B,C) Rabi flopping on axial (B) carrier and (C) first order heating sideband
    before (red) and after (blue) cooling.
    Solid lines in (B) and (C) are theoretical calculation of the Rabi flopping.
    The blue and green lines corresponds to a ground state probability of $92\%$ after cooling and
    the red lines corresponds to a thermal distribution of $70$ $\mu$K before cooling.
    The blue lines does not take into account the effect of decoherence due to resonance
    fluctuation. By comparing it to the green line in (C), which includes a $3$ kHz fluctuation,
    we can clearly see that this effect is the strongest for the after cooling data on
    the axial heating sideband where the Rabi frequency is the lowest.
    \label{f-axial}}
\end{figure*}

@@

Our final cooling results are shown in Fig.~\ref{f-radial} and \ref{f-axial}.
In total, $1000$ cooling pulses (for $100$ ms) are applied along three axes with
cooling begins on the radial second order to remove most energy initially.
To characterize the single atom thermal occupation before and after cooling,
we perform Raman sideband thermometry before and after cooling\fxnote{add citation?}.
For the more tightly confined radial directions (non-degenerate),
we observe clear first-order heating, first-order cooling,
and second-order cooling sidebands before RSC as is shown in red in Figure \ref{f-radial}A.

After cooling, the first and second order cooling sidebands on both radial axes are suppressed.
It is important to note that the simple sideband thermometry formula,
which predicts the ratio of the cooling and heating sideband heights to be $\bar n / (\bar n + 1)$,
assumes that the coupling strength for different motional states to be proportional to $\sqrt{n}$.
Outside of the Lamb-Dickes regime, however,
the coupling strength deviates from this simple scaling rule rapidly
as shown in figure \ref{f-ld}A and since the Raman spectroscopy is performed
with a relatively large Raman Lamb-Dickes parameters $\eta^R$, we cannot use this formula to
measure $\bar n$. Instead, given the absence of the second order cooling sideband,
we assume that the atoms are colder than

\fxnote{How is the thermometry done? pulse duration etc.}
, we perform sideband thermometry
by taking the ratio of the peaks of first-order cooling and heating sidebands to measure $\bar n / (\bar n + 1)$ \fxnote{cite}.
We found $\{\bar n =,\bar n =\}$, which corresponds to a ground state population of $90(2)\%$
and $94(3)\%$ along the two radial directions.

we therefore use an independent measurement of Rabi flopping on the carrier (Fig. \ref{f-radial}B)
and first-order heating(Fig. \ref{f-radial}C) sidebands
to verify the ground-state population~\cite{Meekhof1996}.
The data is fitted using the same Monte-Carlo simulation we used to simulate the cooling
which yields $89(2)\%$ and $92(2)\%$ and show good agreement with
the population extracted from the sideband thermometry.
We also extracted the initial temperature before cooling
from the motional-state decohered Rabi flopping and obtained \fxnote{70(?)} $\mu$K.
